### 2020-12-04 report/article.tex
🎉 Add basic setup with Austin's subfile.

\documentclass[11pt]{article}
\usepackage{times}
\usepackage{setspace}
\singlespacing
\usepackage{amsmath,amssymb, amsthm}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[hang, flushmargin]{footmisc}
\usepackage[colorlinks=true]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{footnotebackref}
\usepackage{url}
\usepackage{listings}
\usepackage[most]{tcolorbox}
\usepackage{inconsolata}
\usepackage[papersize={8.5in,11in}, margin=1in]{geometry}
\usepackage{float}
\usepackage{caption}
\usepackage{esint}
\usepackage{url}
\usepackage{enumitem}
\usepackage{subfig}
\usepackage{wasysym}
\newcommand{\ilcode}{\texttt}
\newcommand{\p}{\partial}
\newcommand{\vphi}{\varphi}
\usepackage{etoolbox}
\usepackage{physics}
\usepackage{xcolor}
\usepackage{subfiles}
\patchcmd{\thebibliography}{\section*{\refname}}{}{}{}

\makeatletter
\renewcommand{\@seccntformat}[1]{}
\makeatother

\begin{document}

\title{\textbf{CSDS 440 Class Project: \\Adversarial Machine Learning}}

\author{Shaochen (Henry) Zhong, \ilcode{sxz517} \\ Minyang Tie, \ilcode{mxt497} \\ Alex Useloff, \ilcode{adu3} \\ Austin Keppers, \ilcode{agk51} \\ David Meshnick, \ilcode{dcm101}\\}
\date{Due and submitted on 12/04/2020 \\ Fall 2020, Dr. Ray}
\maketitle

\section{Introduction}

\section{Shaochen (Henry) Zhong Individual Report}

\section{Minyang Tie Individual Report}

\section{Alex Useloff Individual Report}

\section{Austin Keppers Individual Report}
\subfile{individual_reports/agk51}

\section{David Meshnick Individual Report}

\section{Algorithm Comparison}

\end{document}